---
title: "Choosing a Model"
subtitle: "Demo for deck 02: Programming with LLMs"
format: html
---

Instructor demo: send the same prompt to several providers and compare the
answers, latency, and token usage. Switching providers is a one-string change.

Participants only have Anthropic set up, so this runs as a demo. To compare
providers you need an API key for each one in your `.Renviron`
(`ANTHROPIC_API_KEY`, `OPENAI_API_KEY`, `GOOGLE_API_KEY`, ...). Comment out any
provider you have not configured; Ollama needs a local server running.

Run this interactively (Positron or RStudio), one chunk at a time, rather than
rendering the whole document.

```{r}
#| label: setup
#| message: false
library(ellmer)
library(tidyverse)
```

## List available models

Each provider has a `models_*()` helper that lists its available models. In
Positron, `View()` gives a sortable table.

```{r}
models_anthropic()
```

```{r}
models_openai()
```

```{r}
models_google_gemini()
```

## One prompt, many providers

Switch providers by changing **one string**, of the form `"provider/model"`.

```{r}
prompt <- "Write a recipe for an easy weeknight dinner my kids would like."
```

```{r}
chat("anthropic/claude-sonnet-4-6")$chat(prompt)
```

```{r}
chat("openai/gpt-5")$chat(prompt)
```

```{r}
chat("google_gemini/gemini-2.5-flash")$chat(prompt)
```

## Compare latency and token usage

`ellmer` records token counts on each assistant turn; we time the call ourselves
for latency.

```{r}
compare_model <- function(model, prompt) {
  ch <- chat(model, echo = "none")
  elapsed <- system.time(ch$chat(prompt, echo = "none"))[["elapsed"]]
  tokens <- ch$last_turn()@tokens
  tibble(
    model = model,
    seconds = round(elapsed, 1),
    input_tokens = tokens[[1]],
    output_tokens = tokens[[2]]
  )
}

models <- c(
  "anthropic/claude-haiku-4-5",
  "anthropic/claude-sonnet-4-6",
  "anthropic/claude-opus-4-8",
  "openai/gpt-5",
  "google_gemini/gemini-2.5-flash"
)

results <-
  models |>
  map(compare_model, prompt = prompt) |>
  list_rbind() |>
  arrange(seconds)

results
```
